---
title: "Standard Error Distribution"
author: "Jimmy Zhang"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r packages, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
library(dplyr)
library(tidyr)
library(forcats)
library(readr)
library(ggplot2)
library(rlang)
```

```{r import data}
# Read data
dat_dir <- here("Sim_Data", "CorrectedSE_11262024_Raw")

# List all files matching the pattern
files <- list.files(path = dat_dir, pattern = "results-row-.*\\.rds", full.names = TRUE)
dat_list <- files[order(as.numeric(gsub(".*results-row-(\\d+)\\.rds", "\\1", files)))]

results_list <- lapply(dat_list, function(file) {
  data <- readRDS(file)
})

data_list <- lapply(dat_list, function(file) {
  data <- readRDS(file)
  data[["results"]]  
})

condition_list <- do.call(rbind, lapply(dat_list, function(file) {
  data <- readRDS(file)
  data[["condition"]]  
}))
```

```{r helper function, include=FALSE}
# Examine SE
extract_columns <- function(data, column_names, suffix) {
  columns <- setNames(
    lapply(column_names, function(name) {
      col_name <- paste0(name, suffix)
      if (col_name %in% names(data)) {
        return(data[[col_name]])
      } else {
        warning(paste("Column", col_name, "not found in the data. Returning NA."))
        return(rep(NA, nrow(data)))
      }
    }),
    paste0(column_names, suffix)
  )
  data.frame(columns)
}

# Remove NA
remove_na_columns <- function(df_list) {
  lapply(df_list, function(df) {
    # Remove columns with all NA values
    df[, colSums(is.na(df)) < nrow(df)]
  })
}

# Detect outliers
detect_outliers <- function(column, remove = FALSE) {
  Q1 <- quantile(column, 0.25, na.rm = TRUE)
  Q3 <- quantile(column, 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  outliers <- column[column < lower_bound | column > upper_bound]
  percentage <- length(outliers) / sum(!is.na(column)) * 100
  
  if (remove) {
    column <- ifelse(column < lower_bound | column > upper_bound, NA, column)
  }
  
  list(
    cleaned_column = column,  # Column after removing outliers if remove = TRUE
    outliers = outliers,
    percentage = percentage,
    lower_bound = lower_bound,
    upper_bound = upper_bound
  )
}
```

```{r extract se, include=FALSE}
column_names <- c("joint", "gsam", "lsam", "tspa", "rel")
est_list <- lapply(data_list, extract_columns, column_names = column_names, suffix = ".est")
se_list <- lapply(data_list, extract_columns, column_names = column_names, suffix = ".se")
se_corrected_list <- lapply(data_list, extract_columns, column_names = column_names, suffix = ".se_corrected")
se_corrected_list <- remove_na_columns(se_corrected_list)
```

Standard error distributions are examined under conditions with small sample size (i.e., 30, 60, 120) with low reliability ($\rho = 0.7$).

## Condition 1: $\beta = 0$, $\textit{N}$ = 30, $\rho = 0.7$

```{r conditional data, include=FALSE}
est <- cbind(est_list[[1]], est_list[[1]][c("tspa.est", "rel.est")])
se <- cbind(se_list[[1]], se_corrected_list[[1]])
```

### Density plot of SE distribution (Without removing outliers)

```{r se density plot, echo=FALSE}
# SE Density Plot
par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))
for (i in 1:ncol(se)) {
  emp_sd <- sd(est[[i]], na.rm = TRUE)
  mean_se <- mean(se[[i]], na.rm = TRUE)  # Calculate the mean

  plot(density(se[[i]], na.rm = TRUE), 
       main = paste("Density Plot of SE -", colnames(se)[i]), 
       xlab = "SE", 
       col = "blue", 
       lwd = 2,
       cex.main = 1)     
  
  axis(1, at = seq(0.1, 0.7, by = 0.1))
  
  abline(v = emp_sd, col = "red", lwd = 2, lty = 2)
  abline(v = mean_se, col = "green", lwd = 2, lty = 2)
}
par(mfrow = c(1, 1))
```

### Boxplot of SE values

```{r boxplot, echo=FALSE}
par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))
for (i in 1:ncol(se)) {
  boxplot(se[[i]], 
          main = paste("Boxplot of SE -", colnames(se)[i]), 
          ylab = "SE", 
          col = "lightgreen",
          ylim = c(0, 0.25))
 points(1, mean_se, col = "red", pch = 18, cex = 1.5)
}
par(mfrow = c(1, 1))
```

### Outlier proportions

```{r Detect outliers, echo=FALSE}
outlier_results <- lapply(se, detect_outliers, remove = TRUE)
cleaned_se <- as.data.frame(lapply(outlier_results, `[[`, "cleaned_column"))
outlier_percentages <- sapply(outlier_results, function(res) res$percentage)
outlier_table <- data.frame(
  Method = names(outlier_results),
  Outlier_Percentage = sapply(outlier_results, function(res) res$percentage)
)
print(outlier_table)
```

### Density plot of SE distribution (Removed outliers)

```{r trimmed density plot, echo=FALSE}
# Set up a 3x3 layout for plotting multiple density plots
par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))  # Adjust layout and margins

for (i in 1:ncol(cleaned_se)) {
  emp_sd <- sd(est[[i]], na.rm = TRUE)  # Calculate empirical standard deviation
  mean_se <- mean(cleaned_se[[i]], na.rm = TRUE)  # Calculate the mean of SE
  
  # Plot the density
  plot(
    density(cleaned_se[[i]], na.rm = TRUE), 
    main = paste("Density Plot of SE -", colnames(cleaned_se)[i]), 
    xlab = "SE", 
    col = "blue", 
    lwd = 2,
    xlim = c(0.15, 0.3),  
    ylim = c(0, 20),
    cex.main = 1,  # Slightly larger title font size
    cex.lab = 1.1,   # Increase axis label size
    cex.axis = 0.9   # Slightly smaller axis tick font size
  )
  
  # Add a custom x-axis with formatted tick labels
  axis(
    1, 
    at = seq(0.15, 0.3, by = 0.05), 
    labels = sprintf("%.2f", seq(0.15, 0.3, by = 0.05)), 
    cex.axis = 0.9
  )
  
  # Add vertical lines for empirical SD and mean
  abline(v = emp_sd, col = "red", lwd = 2, lty = 2)  # Red dashed line for SD
  abline(v = mean_se, col = "green", lwd = 2, lty = 2)  # Green dashed line for mean
}

# Reset plotting layout to default
par(mfrow = c(1, 1))
```



## Condition 2: $\beta = 0.3$, $\textit{N}$ = 30, $\rho = 0.7$

### Density plot of SE distribution (Without removing outliers)

```{r conditional data 2}
est <- cbind(est_list[[3]], est_list[[3]][c("tspa.est", "rel.est")])
se <- cbind(se_list[[3]], se_corrected_list[[3]])
```

```{r se density plot 2}
# SE Density Plot
par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))
for (i in 1:ncol(se)) {
  emp_sd <- sd(est[[i]], na.rm = TRUE)
  mean_se <- mean(se[[i]], na.rm = TRUE)  # Calculate the mean

  plot(density(se[[i]], na.rm = TRUE), 
       main = paste("Density Plot of SE -", colnames(se)[i]), 
       xlab = "SE", 
       col = "blue", 
       lwd = 2,
       cex.main = 1)     
  
  axis(1, at = seq(0.1, 0.7, by = 0.1))
  
  abline(v = emp_sd, col = "red", lwd = 2, lty = 2)
  abline(v = mean_se, col = "green", lwd = 2, lty = 2)
}
par(mfrow = c(1, 1))
```

### Boxplot of SE values

```{r boxplot 2, echo=FALSE}
par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))
for (i in 1:ncol(se)) {
  boxplot(se[[i]], 
          main = paste("Boxplot of SE -", colnames(se)[i]), 
          ylab = "SE", 
          col = "lightgreen",
          ylim = c(0, 0.25))
 points(1, mean_se, col = "red", pch = 18, cex = 1.5)
}
par(mfrow = c(1, 1))
```

### Outlier proportions

```{r Detect outliers 2}
outlier_results <- lapply(se, detect_outliers, remove = TRUE)
cleaned_se <- as.data.frame(lapply(outlier_results, `[[`, "cleaned_column"))
outlier_percentages <- sapply(outlier_results, function(res) res$percentage)
outlier_table <- data.frame(
  Method = names(outlier_results),
  Outlier_Percentage = sapply(outlier_results, function(res) res$percentage)
)
print(outlier_table)
```

### Density plot of SE distribution (Removed outliers)

```{r trimmed density plot 2, echo=FALSE}
# Set up a 3x3 layout for plotting multiple density plots
par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))  # Adjust layout and margins

for (i in 1:ncol(cleaned_se)) {
  emp_sd <- sd(est[[i]], na.rm = TRUE)  # Calculate empirical standard deviation
  mean_se <- mean(cleaned_se[[i]], na.rm = TRUE)  # Calculate the mean of SE
  
  # Plot the density
  plot(
    density(cleaned_se[[i]], na.rm = TRUE), 
    main = paste("Density Plot of SE -", colnames(cleaned_se)[i]), 
    xlab = "SE", 
    col = "blue", 
    lwd = 2,
    xlim = c(0.15, 0.4),  
    ylim = c(0, 20),
    cex.main = 1,  # Slightly larger title font size
    cex.lab = 1.1,   # Increase axis label size
    cex.axis = 0.9   # Slightly smaller axis tick font size
  )
  
  # Add a custom x-axis with formatted tick labels
  axis(
    1, 
    at = seq(0.15, 0.3, by = 0.05), 
    labels = sprintf("%.2f", seq(0.15, 0.3, by = 0.05)), 
    cex.axis = 0.9
  )
  
  # Add vertical lines for empirical SD and mean
  abline(v = emp_sd, col = "red", lwd = 2, lty = 2)  # Red dashed line for SD
  abline(v = mean_se, col = "green", lwd = 2, lty = 2)  # Green dashed line for mean
}

# Reset plotting layout to default
par(mfrow = c(1, 1))
```


## Condition 3: $\beta = 0.6$, $\textit{N}$ = 30, $\rho = 0.7$

### Density plot of estimates distribution

```{r conditional data 3}
est <- cbind(est_list[[5]], est_list[[5]][c("tspa.est", "rel.est")])
se <- cbind(se_list[[5]], se_corrected_list[[5]])
```

```{r est density plot, echo=FALSE}
# Estimates Density Plot
# Setting up the plotting area for a 3x3 grid layout
par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))

# Calculate density for each column to determine consistent y-axis limits
density_list <- lapply(1:ncol(est), function(i) density(est[[i]], na.rm = TRUE))
max_y <- max(sapply(density_list, function(d) max(d$y)))

# Loop through each column of 'est' to create density plots
for (i in 1:ncol(est)) {
  emp_mean <- mean(est[[i]], na.rm = TRUE)
  
  # Plot the density for each column with a consistent y-axis limit
  plot(density_list[[i]], 
       main = paste("Density Plot of SE -", colnames(est)[i]), 
       xlab = "SE", 
       col = "blue", 
       lwd = 2,
       xlim = c(-1, 1),  
       ylim = c(0, max_y),  # Consistent y-axis across all plots
       cex.main = 1)     
  
  # Add a green dashed line at the empirical mean
  abline(v = emp_mean, col = "green", lwd = 2, lty = 2)
  
  # Add a vertical reference line at zero
  abline(v = 0.6, col = "red", lwd = 2, lty = 2)
}

# Reset the plotting area to a single plot
par(mfrow = c(1, 1))
```

### Density plot of SE distribution (Without removing outliers)

```{r se density plot 3}
# SE Density Plot
par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))
for (i in 1:ncol(se)) {
  emp_sd <- sd(est[[i]], na.rm = TRUE)
  mean_se <- mean(se[[i]], na.rm = TRUE)  # Calculate the mean

  plot(density(se[[i]], na.rm = TRUE), 
       main = paste("Density Plot of SE -", colnames(se)[i]), 
       xlab = "SE", 
       col = "blue", 
       lwd = 2,
       cex.main = 1)     
  
  axis(1, at = seq(0.1, 0.7, by = 0.1))
  
  abline(v = emp_sd, col = "red", lwd = 2, lty = 2)
  abline(v = mean_se, col = "green", lwd = 2, lty = 2)
}
par(mfrow = c(1, 1))
```

### Boxplot of SE values

```{r boxplot 3, echo=FALSE}
par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))
for (i in 1:ncol(se)) {
  boxplot(se[[i]], 
          main = paste("Boxplot of SE -", colnames(se)[i]), 
          ylab = "SE", 
          col = "lightgreen",
          ylim = c(0, 0.25))
 points(1, mean_se, col = "red", pch = 18, cex = 1.5)
}
par(mfrow = c(1, 1))
```

### Outlier proportions

```{r Detect outliers 3}
outlier_results <- lapply(se, detect_outliers, remove = TRUE)
cleaned_se <- as.data.frame(lapply(outlier_results, `[[`, "cleaned_column"))
outlier_percentages <- sapply(outlier_results, function(res) res$percentage)
outlier_table <- data.frame(
  Method = names(outlier_results),
  Outlier_Percentage = sapply(outlier_results, function(res) res$percentage)
)
print(outlier_table)
```

### Density plot of SE distribution (Removed outliers)

```{r trimmed density plot 3, echo=FALSE}
# Set up a 3x3 layout for plotting multiple density plots
par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))  # Adjust layout and margins

for (i in 1:ncol(cleaned_se)) {
  emp_sd <- sd(est[[i]], na.rm = TRUE)  # Calculate empirical standard deviation
  mean_se <- mean(cleaned_se[[i]], na.rm = TRUE)  # Calculate the mean of SE
  
  # Plot the density
  plot(
    density(cleaned_se[[i]], na.rm = TRUE), 
    main = paste("Density Plot of SE -", colnames(cleaned_se)[i]), 
    xlab = "SE", 
    col = "blue", 
    lwd = 2,
    xlim = c(0, 0.4),  
    ylim = c(0, 20),
    cex.main = 1,  # Slightly larger title font size
    cex.lab = 1.1,   # Increase axis label size
    cex.axis = 0.9   # Slightly smaller axis tick font size
  )
  
  # Add a custom x-axis with formatted tick labels
  axis(
    1, 
    at = seq(0.15, 0.3, by = 0.05), 
    labels = sprintf("%.2f", seq(0.15, 0.3, by = 0.05)), 
    cex.axis = 0.9
  )
  
  # Add vertical lines for empirical SD and mean
  abline(v = emp_sd, col = "red", lwd = 2, lty = 2)  # Red dashed line for SD
  abline(v = mean_se, col = "green", lwd = 2, lty = 2)  # Green dashed line for mean
}

# Reset plotting layout to default
par(mfrow = c(1, 1))
```

### Confidence intervals (Without removing outliers)

```{r CI without outliers, echo=FALSE, message=FALSE, warning=FALSE}
est <- est[1:300, ]
se <- se[1:300, ]

n_methods <- ncol(est)  # Number of methods
for (i in 1:n_methods) {

  estimates <- est[, i]
  se_values <- se[, i]
  
  lower_bound <- estimates - 1.96 * se_values
  upper_bound <- estimates + 1.96 * se_values
  
  data <- data.frame(
    Replication = 1:300,
    Estimate = estimates,
    Lower = lower_bound,
    Upper = upper_bound
  )
  
  plot(
    data$Replication, data$Estimate, 
    pch = 16, col = "blue", ylim = c(-2, 2), 
    xlab = "Replication", ylab = "Estimate", 
    main = paste("Confidence Intervals -", colnames(est)[i]),
    cex = 0.8
  )
  
  arrows(
    x0 = data$Replication, y0 = data$Lower,
    x1 = data$Replication, y1 = data$Upper,
    code = 3, angle = 90, length = 0.05, col = "black"
  )
  
  abline(h = 0.6, col = "red", lty = 2, lwd = 2)
}

```

\newpage

The green line represents the mean of estimated standard errors, while the red line is the empirical standard deviation of path coefficients. The green line (mean standard error) is consistently positioned to the left of the red line (empirical standard deviation), indicating that the model underestimates the true standard error. 

However, the magnitude of this underestimation varies between methods:

- gsam.se and lsam.se: The difference between the red and green lines is relatively small compared to other methods, indicating that these methods might provide a closer estimate of the true variability.
- rel.se and tspa.se_corrected: There is a more substantial difference between the red and green lines, suggesting that these methods tend to underestimate the standard error more compared to others.

For shape and spread of density curves:

- The density curves for joint.se, gsam.se, lsam.se, and tspa.se are relatively similar, with a peak around 0.2 to 0.3 and a sharp decline, indicating that these methods have similar distributions of standard errors.
- The tspa.se_corrected and rel.se methods show a slightly broader spread, suggesting greater variability in the calculated standard errors.








