---
title: "tspa-corrected-se"
author: "Jimmy Zhang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tspa-interaction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r load packages, include=FALSE}
library(lavaan)
library(dplyr)
library(here)
library(MASS)
r_scripts <- list.files(here("R"), pattern = "\\.R$", full.names = TRUE)
lapply(r_scripts, source)
```

```{r data-generating function, echo=FALSE}
sim_sem_data <- function(n_obs, lambda, phi, theta, beta = NULL, psi = NULL) {
  n_factors <- ncol(lambda)
  n_indicators <- nrow(lambda)
  if (!is.null(beta) && !is.null(psi)) {
    phi <- solve(diag(n_factors) - beta) %*% psi %*% t(solve(diag(n_factors) - beta))
  }
  sigma <- lambda %*% phi %*% t(lambda) + diag(theta)
  data <- mvrnorm(n = n_obs, mu = rep(0, n_indicators), Sigma = sigma)
  
  data <- as.data.frame(data)
  colnames(data) <- paste0("y", 1:n_indicators)
  return(data)
}
```

```{r simulate data, echo=TRUE}
set.seed(4443)
lambda <- matrix(c(0.7, 0.8, 0.9, 0, 0, 0,
                   0, 0, 0, 0.6, 0.7, 0.8), 
                 nrow = 6, ncol = 2)

psi <- matrix(c(1, 0, 
                0, 1), nrow = 2, ncol = 2)

# Define path coefficients between latent factors (e.g., F2 ~ F1)
beta <- matrix(c(0, 0.8, 
                 0, 0), nrow = 2, ncol = 2) 

theta <- c(1.3, 1.2, 1.1, 1.25, 1.15, 1.05)

df <- sim_sem_data(n_obs = 500, lambda = lambda, phi = NULL, theta = theta, beta = beta, psi = psi)

# Test the sim data
mod <- "f1 =~ y1 + y2 + y3
        f2 =~ y4 + y5 + y6
        f2 ~ f1"
summary(sem(mod, df, std.lv = T), standardized = TRUE)
```

```{r an example for Bartlett scores}
# Factor score
tspa_bt <- get_fs(df,
                  model = "f1 =~ y1 + y2 + y3
                           f2 =~ y4 + y5 + y6",
                  method = "Bartlett",
                  vfsLT = TRUE)
vfsLT_bt <- attr(tspa_bt, which = "vfsLT") # Sampling covariance matrix of factor score loading and error variance

tspa_bt_fit <- tspa(model = "f2 ~ f1",
                    data = tspa_bt,
                    fsT = attr(tspa_bt, "fsT"), # Error variance
                    fsL = diag(2) |>
                       `dimnames<-`(list(c("fs_f1", "fs_f2"), # Factor loadings fixed at 1 for Bartlett factor scores
                                         c("f1", "f2"))))

# Adjusted covariance matrix
corrected_vcov <- vcov_corrected(
                      tspa_bt_fit,
                      # Exclude fixed loadings and error variance
                      vfsLT = vfsLT_bt[c(5, 7), c(5, 7)],
                      # Specify which elements are free (error variances only)
                      which_free = c(5, 7))

# Unadjusted standard errors
sqrt(diag(vcov(tspa_bt_fit)))
# Corrected standard errors
sqrt(diag(corrected_vcov))
```

This `vcov_corrected function` (by Mark) is designed to correct the covariance matrix of parameter estimates in a 2S-PA model. It accounts for the sampling covariance of the factor score loadings (`fsL`) and the factor score error covariances (`fsT`). This correction is crucial because it adjusts for the additional variance introduced when using estimated factor scores in the structural model. 

```{r decomposition of `vcov_corrected function`, eval=FALSE, include=TRUE}
# The function extracts the lower triangular part of each fsT matrix (including the diagonal) and stores these elements in a list val_fsT_lower.
# It then combines (unlists) fsL and the lower triangular parts of fsT into a single vector val_fsLT, which contains all the parameters of interest for correction.

fsL = diag(2) |>
  `dimnames<-`(list(c("fs_f1", "fs_f2"),
                    c("f1", "f2")))
fsT = attr(tspa_bt, "fsT")

val_fsT_lower <- lapply(fsT, function(x) {
        x[lower.tri(x, diag = TRUE)]
    })
val_fsLT <- c(unlist(fsL), unlist(fsT))

# Determining Which Parameters Are Free
which_free <- c(5, 7)

# Jacobian Matrix Calculation
# The function calculates the Jacobian matrix (jac) using lavaan::lav_func_jacobian_complex. This matrix captures the partial derivatives of the model parameters with respect to the free parameters in fsL and fsT.
jac <- lavaan::lav_func_jacobian_complex(
        function(x) {
            par <- val_fsLT
            par[which_free] <- x
            counter <- 0
            num_ld <- length(val_fsL[[1]])
            for (g in seq_len(ngrp)) {
                val_fsL[[g]][] <- par[(1:num_ld) + counter]
                counter <- counter + num_ld
            }
            num_ev <- sum(lower.tri(val_fsT[[1]], diag = TRUE))
            for (g in seq_len(ngrp)) {
                val_fsT[[g]][] <- lavaan::lav_matrix_lower2full(
                    par[(1:num_ev) + counter]
                )
                counter <- counter + num_ev
            }
            if (ngrp == 1) {
                val_fsL <- val_fsL[[1]]
                val_fsT <- val_fsT[[1]]
            }
            lavaan::coef(update_tspa(tspa_fit, fsL = val_fsL, fsT = val_fsT))
        },
        x = val_fsLT[which_free],
    )

# Computing the Corrected Covariance Matrix
vcov(tspa_fit) + jac %*% vfsLT %*% t(jac)
```



